% Please number citations consecutively within brackets \cite{b1}. The
% sentence punctuation follows the bracket \cite{b2}. Refer simply to the reference
% number, as in \cite{b3}---do not use ``Ref. \cite{b3}'' or ``reference \cite{b3}'' except at
% the beginning of a sentence: ``Reference \cite{b3} was the first $\ldots$''

% Capitalize only the first word in a paper title, except for proper nouns and element
% symbols.

\newpage


\begin{thebibliography}{00}
\bibitem{stonebraker}
  M. Stonebraker, S. Madden, D. J. Abadi,  S. Harizopoulos, N. Hachem, \& P. Helland,
  ``The end of an architectural era: It's time for a complete rewrite,''
  Proceedings of the VLDB Endowment,
  pp. 1150--1160,
  2007.
\bibitem{bernstein}
  P. A. Bernstein, V. Hadzilacos, \& N. Goodman,
  ``Concurrency control and recovery in database systems'',
  Reading: Addison-wesley,
  vol. 370,
  1987.
\bibitem{guerraoui}
  R. Guerraoui and J. Wang,
  ``How fast can a distributed transaction commit?'',
  Proceedings of the 36th ACM SIGMOD-SIGACT-SIGAI Symposium on Principles of
  Database System,
  pp. 107--122,
  May 2017.
\bibitem{harding}
  R. Harding, D. Van Aken, A. Pavlo, \& M. Stonebraker,
  ``An evaluation of distributed concurrency control'',
  Proceedings of the VLDB Endowment,
  vol. 10(5),
  pp. 553--564,
  January 2017.
\bibitem{curino}
  C. Curino, Y. Zhang, E. P. C. Jones \& S. Madden,
  ``Schism: a workload-driven approach to database replication and partitioning'',
  Proceedings of the VLDB Endowment,
  vol. 3(1),
  pp. 48--57,
  2010.
\bibitem{das}
  S. Das, D. Agrawal \& A. El Abbadi,
  ``G-Store: a scalable data store for transactional multi key access in the cloud'',
  Proceedings of the 1st {ACM} Symposium on Cloud Computing,
  pp. 163--174,
  June 2010.
\bibitem{lin}
  Q. Lin, P. Chang, G. Chen, B. C. Ooi,  K. L. Tan \& Z. Wang,
  ``Towards a non-2PC transaction management in distributed database systems'',
  Proceedings of the {ACM} {SIGMOD} International Conference on Management of
  Data,
  pp. 1659--1674,
  2016.
\bibitem{thomson}
  A. Thomson, T. Diamond, S. Weng, K. Ren, P. Shao \& D. J. Abadi,
  ``Calvin: fast distributed transactions for partitioned database systems'',
  Proceedings of the {ACM} {SIGMOD} International Conference on Management of
  Data,
  pp. 1--12,
  2012
\bibitem{ren}
  K. Ren, D. Li \& D. J. Abadi,
  ``SLOG: serializable, low-latency, geo-replicated transactions''
  Proceedings of the VLDB Endowment,
  vol. 12(11),
  pp. 1747--1761,
  2019.
\bibitem{lu}
  Y. Lu, X. Yu, L. Cao \& S. Madden,
  ``Epoch-based commit and replication in distributed {OLTP} databases'',
  Proceedings of the VLDB Endowment,
  vol. 14(5),
  pp. 743--756,
  2021.
  \bibitem{dewitt}
  D. J. DeWitt, Randy H. Katz, Frank Olken, L. D. Shapiro, M. Stonebraker
  \&  D. A. Wood,
  ``Implementation Techniques for Main Memory Database Systems'',
  SIGMOD'84, Proceedings of Annual Meeting,
  pp. 1--8,
  1984.
  \bibitem{crooks}
  N. Crooks, M. Burke, E. Cecchetti, S. Harel, R. Agarwal \& L. Alvisi,
  ``Obladi: oblivious serializable transactions in the cloud'',
  13th {USENIX} Symposium on Operating Systems Design and Implementation,
  pp. 727--743,
  2018.
\bibitem{tu}
  S. Tu, W. Zheng, E. Kohler, B. Liskov \& S. Madden,
  ``Speedy transactions in multicore in-memory databases'',
  {ACM} {SIGOPS} 24th Symposium on Operating Systems Principles,
  pp. 18--32,
  2013.
\bibitem{narula}
  N. Narula, C. Cutler, E. Kohler \& R. T. Morris,
  ``Phase Reconciliation for Contended In-Memory Transactions'',
  11th {USENIX} Symposium on Operating Systems Design and Implementation,
  pp. 511--524,
  2014.
 \bibitem{ongaro}
  D. Ongaro and J. K. Ousterhout,
  ``In Search of an Understandable Consensus Algorithm'',
  {USENIX} Annual Technical Conference,
  pp. 305--319,
  2014.
\bibitem{lamport}
  L. Lamport,
  ``The part-time parliament'',
  {ACM} Trans. Comput. Syst.,
  vol. 16(2),
  pp. 133--169,
  1998.
  \bibitem{garraghan}
  P. Garraghan, P. Townend, \& J. Xu,
  ``An empirical failure-analysis of a large-scale cloud computing environment'',
  IEEE 15th International Symposium on High-Assurance Systems Engineering,
  pp. 113--120,
  January 2014.
  \bibitem{birke} 
  R. Birke, I. Giurgiu, L. Y. Chen, D. Wiesmann \& T. Engbersen,
  ``Failure analysis of virtual and physical machines: patterns, causes and characteristics'',
  44th Annual {IEEE/IFIP} International Conference on Dependable Systems and Networks,
  pp. 1--12,
  2014.
  \bibitem{wang}
  G. Wang, L. Zhang \& W. Xu,
  ``What can we learn from four years of data center hardware failures?``,
  47th Annual {IEEE/IFIP} International Conference on Dependable Systems and Networks, 
  pp. 25--36,
  2017.  
 \bibitem{bailey}
  N.T.J. Bailey,
  ``On Queueing Processes with Bulk Service'',
  Journal of the Royal Statistical Society, Series B 16, 1,
  pp. 80--87,
  1954.
  \bibitem{tpcc}
  TPC-C Benchmark C
  ``https://www.tpc.org/tpcc/'',
  2010.
\bibitem{tarjan}
  R. Tarjan,
  ``Depth-first search and linear graph algorithms'',
  SIAM journal on computing,
  vol. 1(2),
  pp.146--160,
  1972.  
  \bibitem{mitrani}
  I. Mitrani,
  ``Simulation techniques for discrete event systems'',
  Cambridge University Press,
  1982.
  \bibitem{thomson2}
  A. Thomson and D. J. Abadi,
  ``The case for determinism in database systems'',
  Proceedings of the VLDB Endowment,
  vol. 3(1),
  pp. 70--80,
  2010.
\bibitem{ren2}
  K. Ren, A. Thomson \& D. J. Abadi,
  ``An evaluation of the advantages and disadvantages of deterministic database
  systems'',
  Proceedings of the VLDB Endowment,
  vol. 7(10),
  pp. 821--832,
  2014.
\bibitem{lu2}
  Y. Lu, X. Yu, L. Cao \& S. Madden,
  ``Aria: {a} fast and practical deterministic {OLTP} database'',
  Proceedings of the VLDB Endowment,
  vol. 13(11),
  pp. 2047--2060,
  2020.
\bibitem{issa}
  S. Issa, M. Viegas, P. Raminhas, N. Machado, M. Matos \& P. Romano,
  ``Exploiting symbolic execution to accelerate deterministic databases'',
  40th {IEEE} International Conference on Distributed Computing Systems,
  pp. 678--688,
  2020.  
\bibitem{du}
  J. Du, S. Elnikety \& W. Zwaenepoel,
  ``Clock-{SI}: snapshot isolation for partitioned data stores using loosely synchronized clocks'',
  {IEEE} 32nd Symposium on Reliable Distributed Systems,
  pp. 173--184,
  2013.
\bibitem{elnikety} 
  S. Elnikety, W. Zwaenepoel \& F. Pedone,
  ``Database Replication Using Generalized Snapshot Isolation'',
  24th {IEEE} Symposium on Reliable Distributed Systems,
  pp. 77--84,
  2005.
 \bibitem{sovran}
  Y. Sovran, R. Power, M. K.. Aguilera \& J. Li,
  ``Transactional storage for geo-replicated systems'',
  Proceedings of the 23rd {ACM} Symposium on Operating Systems Principles,
  pp. 385--400,
  2011.
\bibitem{bailis}
  P. Bailis, A. Davidson, A. Fekete, A. Ghodsi, J. M. Hellerstein, \& I. Stoica,
  ``Highly available transactions: virtues and limitations'',
  Proceedings of the VLDB Endowment,
  vol. 7(3),
  pp. 181--192,
  2013.
\bibitem{kraska}
  T. Kraska, G. Pang, M. J. Franklin, S. Madden \& A. D. Fekete,
  ``{MDCC:} multi-data center consistency'',
  Eighth Eurosys Conference,
  pp. 113--126,
  2013.
\bibitem{zhang}
  I. Zhang, N. K. Sharma, A. Szekeres, A. Krishnamurthy \& D. R. K. Ports,
  ``Building consistent transactions with inconsistent replication'',
  Proceedings of the 25th Symposium on Operating Systems Principles,
  pp. 263--287,
  2015.
\bibitem{fan}
  H. Fan and W. M. Golab,
  ``Ocean vista: gossip-based visibility control for speedy geo-distributed transactions'',
  Proceedings of the VLDB Endowment,
  vol. 12(11),
  pp. 1471--1484,
  2019.
\bibitem{mu}
  S. Mu, L. Nelson, W. Lloyd \& J. Li,
  ``Consolidating concurrency control and consensus for commits under conflicts'',
  12th {USENIX} Symposium on Operating Systems Design and Implementation,
  pp. 517--532,
  2016.
\bibitem{nawab}
  F. Nawab, V. Arora, D. Agrawal \& A. El Abbadi,
  ``Minimizing commit latency of transactions in geo-replicated data stores'',
  Proceedings of the 2015 {ACM} {SIGMOD} International Conference on Management
  of Data,
  pp. 1279--1294,
  2015.
\bibitem{maiyya}
  S. Maiyya, F. Nawab, D. Agrawal \& A. El Abbadi,
  ``Unifying consensus and atomic commitment for effective cloud data management''
  Proceedings of the VLDB Endowment,
  vol. 12(15),
  pp. 611--623,
  2019.
\bibitem{taft}
  Taft et al.,
  ``CockroachDB: the resilient geo-distributed {SQL} database'',
  Proceedings of the {ACM} {SIGMOD} International Conference on Management of
  Data,
  pp. 1493--1509,
  2020.
\bibitem{lu3}
  Y. Lu, X. Yu \& S. Madden,
  ``{STAR}: scaling transactions through asymmetric replication'',
  Proceedings of the VLDB Endowment,
  vol. 12(11),
  pp. 1316--1329,
  2019.
  \bibitem{wu}
  H. Wu, Z. Shang and K. Wolter,
  ``Performance Prediction for the Apache Kafka Messaging System'',
  {IEEE} 21st International Conference on High Performance Computing and Communications;
  {IEEE} 17th International Conference on Smart City;
  {IEEE} 5th International Conference on Data Science and Systems,
  pp. 154-161,
  2019.
  \bibitem{meng}
  T. Meng, Y. Zhao, K. Wolter \& C. Xu,
  ``On consortium blockchain consistency: {a} queueing network model approach'',
  {IEEE} Trans. Parallel Distributed Syst.,
  vol. 32(6),
  pp. 1369--1382,
  2021.
  \bibitem{yu}
  Yu, Philip S., D. M. Dias \& S. S. Lavenberg.
  ``On the analytical modeling of database concurrency control'',
  Journal of the ACM,
  vol. 40(4), 
  pp. 831--872,
  1993.
  \bibitem{agrawal}
  R. Agrawal and D. J. DeWitt,
  ``Integrated concurrency control and recovery mechanisms: design and performance evaluation'',
  {ACM} Trans. Database Syst.,
  vol. 10(4),
  pp. 529--564,
  1985.
  \bibitem{disanzo}
  P. di Sanzo, B. Ciciani, R. Palmieri, F. Quaglia and P. Romano,
  ``On the analytical modeling of concurrency control algorithms for software transactional memories: the case of commit-time-locking'',
  Perform. Evaluation,
  vol. 69(5),
  pp. 187--205,
  2012.
  \bibitem{menasce}
  A. Menasce and T. Nakanishi,
  ``Performance evaluation of a two-phase commit based protocol for {DDBS}'',
  Proceedings of the {ACM} Symposium on Principles of Database Systems,
  pp. 247--255,
  1982.  
  \bibitem{liu}
  M.L. Liu, D. Agrawal \& A. El Abbadi,
  ``The performance of two phase commit Protocols in the presence of site failures'',
  Distributed and Parallel Databases,
  vol. 6,
  pp. 157--182,
  1998.

\end{thebibliography}
\vspace{12pt}


